---
title: "Understanding the influence of transportation infrastrucure on Urban Heat in Phoenix, Arizona"
author: "C.G. Hoehne"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}

# default to knitr option of code printing
knitr::opts_chunk$set(echo = TRUE)

# load fonts for use figures
#extrafont::loadfonts(device = "win") 
##** CHANGE THIS TO LOAD ONLY THE FONT WE WILL USE
##** PREFERABLY INCLUDE THIS FONT IN THE REPO

# store master figure font as string
my.font <- "Helvetica"

# list of all dependant packages
list.of.packages <- c("tidyverse",
                      "data.table", 
                      "lubridate",
                      "weathermetrics",
                      "rgdal",
                      "rgeos",
                      "raster",
                      "tmap",
                      "maptools",
                      "deldir",
                      "xtable",
                      "knitr",
                      "tinytex",
                      "here")

# install missing packages
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# install TinyTex (for creating PDFs from rmarkdown) if it is not installed.
# NOTE YOU NEED TO RESTART R STUDIO AFTER INSTALL
#tinytex::install_tinytex(force = FALSE)

# load packages
lapply(list.of.packages, library, character.only = TRUE)

# store loaded packages and thier version in a data.frame
package.info <- data.frame("Package" = list.of.packages, "Version" = unlist(lapply(list.of.packages, function (x) as.character(packageVersion(x)))))

```

## Data Sources

### Weather Data
- National Centers for Environmental Information (NCEI): https://gis.ncdc.noaa.gov/maps/ncei/cdo/hourly
- The Arizona Meteorological Network (AZMET): https://cals.arizona.edu/azmet/az-data.htm
- Maricopa County Flood Control District (MCFCD): https://www.maricopa.gov/3769/Weather-Sensor-Data
- iButton data recorded in the Summer of 2017 by the Urban Water Innovation Network (UWIN): https://erams.com/UWIN/


### Parking Data
- Forthcoming publication by Hoehne et al, info here: http://transportationlca.org/phoenixparking/

### Road Network Data
- OpenStreetMap (© OpenStreetMap contributors)
- Arizona Department of Transportation (ADOT)

### Traffic Data
- Calibrated Maricopa Association of Governments (MAG) travel demand and parcel data via modifed version of SUMO (Simulation of Urban MObility) using OpenStreetMap street network data


```{r data import, include = FALSE}

## import weather, roadway, traffic & parking data

# cleaned uza weather data
uza.weather <- readRDS(here("data/outputs/2017-uza-weather-data.rds"))

# uza station data (spatial points w/ metadata)
uza.stations <- readRDS(here("data/outputs/uza-station-data.rds"))

# formated UZA OSM data (list of spatial objects, each element is different roadway class that is buffered link according to roadway width)
#uza.roads <- readRDS(here("data/outputs/osm-links-buffers-sp-list.rds"))

# cleaned traffic data
# **(temporary dummy traffic data by osm_id and hour of day, e.g. dummy ICARUS output)**
dummy.traffic <- as.data.table(matrix(runif(100, 100, 6000), nrow = 100, ncol = 24) * runif(24, 0, 1))
dummy.traffic <- dummy.traffic[, floor(.SD)] # round down to nearest whole number for whole data.table
colnames(dummy.traffic) <- paste0("hour", seq(1:24)) # rename cols to "hour#" (1:24)
dummy.traffic$osm_id <- seq.int(nrow(dummy.traffic)) # dummy osm_id column (1:100)

# cleaned parking data 
# *** PLACEHOLDER *** 


## import other various shapefiles for plotting
city.labels <- shapefile(here("data/shapefiles/other/phx_metro_labels.shp")) # city labels (pre-cropped to UZA)
uza.border <- shapefile(here("data/shapefiles/boundaries/maricopa_county_uza.shp")) # UZA 
cnty.border <- shapefile(here("data/shapefiles/boundaries/maricopa_county.shp")) # county 
highways <- shapefile(here("data/shapefiles/other/phx_metro_hways.shp")) # 2017 highways (pre-cropped to UZA)

```

## Plots

Here are the working plots:

```{r preliminary data plots, echo = FALSE}

#-#-#-#-#-#-#-#
# Data Plots #
#-#-#-#-#-#-#

# set knitr chunk options to put spaces between plots
opts_chunk$set(fig.ncol = 3, fig.sep = c('\\newline','\\newline','\\newline')) # only workds for LaTex output

# store palette for plotting stations
my.palette <- c("#0C120C","#640D14","#C20114","#6D7275")

# plot station points with UZA, highways, & city labels for context.
stations.plot <-   
  tm_shape(uza.border) + tm_borders(lwd = 0.8, lty = "solid", col = "grey40", alpha = 0.7) + tm_fill(col = "grey90") + # urbanized area (UZA) border & fill
  tm_shape(w.stations.spdf, bbox = w.stations.spdf) + tm_dots(col = "source", border.col = NULL, palette = my.palette, size = 0.2, alpha = 0.8, title = "Data Source") + # station points
  tm_shape(cnty.border) + tm_borders(lwd = 0.8, lty = "solid", col = "grey20", alpha = 0.7) + # county border
  tm_scale_bar(position = c(0.4,0.0), breaks = c(0,5,10,15,20), size = 0.90, color.light = "grey85") + # scalebar
  tm_compass(north = 0, type = "4star", size = 2, show.labels = 1, position = c(0.9,0.85)) + # compass
  tm_shape(highways) + tm_lines(lwd = 1.1, col = "grey40") + # highways
  tm_shape(city.labels) + tm_text("name", size = .5, fontfamily = my.font,  fontface = "bold.italic", just = "center") + # city labels
  tm_layout(fontfamily = my.font, fontface = "italic", bg.color = "grey95", title.size = 1.5, legend.title.size = 1.10, # theme & formatting
            legend.text.size = 0.95, title = "", title.position = c(0.12,0.94),
            legend.position = c(0.02,0.45), outer.margins = c(0,0,0,0), asp = 0)

#tmap_save(stations.plot, filename = here("figures/stations_w_ibut.png")) # option to save
stations.plot #(plot in report)


# calculate range of temps and mean temp for urbanized stations

# first clip stations points to ones w/in uza 
uza.stations <- raster::intersect(w.stations.spdf, uza.border)

# filter station list to uza stations
uza.stations.list <- w.stations[station.name %in% uza.stations$station.name]

# if time is within 15 min of hour, round to hour, otherwise drop, then
# aggregate to the min, mean, max temp at each unique timestep for all stations
uza.weather[, date.time.round := as.POSIXct(round.POSIXt(date.time, "hours"))]
uza.weather <- uza.weather[abs(difftime(date.time, date.time.round, units = "min")) < 15]
uza.weather.agg <- uza.weather[, .(min.temp.f = min(temp.f, na.rm = T),
                             med.temp.f = median(temp.f, na.rm = T),
                             max.temp.f = max(temp.f, na.rm = T)), by = .(date.time.round)]

# hourly temp min/mean/max by 2017 month aggregate across all stations
uza.weather[, month := as.factor(month(date.time.round, label = T))]
uza.weather[, hour := hour(date.time.round)]
uza.weather.month.agg <- uza.weather[, .(min.temp.f = min(temp.f, na.rm = T),
                                   med.temp.f = median(temp.f, na.rm = T),
                                   max.temp.f = max(temp.f, na.rm = T)), by = .(hour,month)]

# plot hourly temp.f range by month for 2017 for all stations
# Plot historical tempeature plots of metro regions
month.x.hour.p <- ggplot(uza.weather.month.agg, aes(x = hour, y = med.temp.f , group = month)) +
  geom_ribbon(aes(ymin = min.temp.f, ymax = max.temp.f),  fill = "grey50", alpha = 0.5) +
  geom_line(size = 1.2) +
  geom_segment(aes(x = 0, y = 25, xend = 23, yend = 25)) +
  geom_segment(aes(x = 0, y = 25, xend = 0, yend = 135)) +
  facet_wrap(~ month , scales = "fixed") +
  labs(title = "Hourly Temperature Ranges by month for Phoenix UZA, 2017", x = "Hour", y = "Temperature (deg F)") +
  scale_x_continuous(limits = c(0,23), breaks = c(0,6,12,18), expand = c(0,0)) +
  scale_y_continuous(limits = c(25,135), breaks = c(25,50,75,100,125), expand = c(0,0)) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", family = my.font, size = 12, colour = "black", hjust = 1),
        strip.text = element_text(family = my.font, size = 11, colour = "black"),
        axis.title = element_text(face = "bold", family = my.font, size = 12, colour = "black"),
        axis.text = element_text(family = my.font, size = 10, colour = "black"),
        axis.text.x = element_text(hjust = 0.25) # for some reason the text isn't defaulted to centered on the ticks
  )

month.x.hour.p
#ggsave("hrly_tempF_by_month_all_phx_UZA_stations.png", month.x.hour.p, device = "png", path = here("figures"), scale = 1, width = 6.5, height = 8, dpi = 300, units = "in")

# plot hourly temp.f range for summer months for 2017 for all stations
uza.weather.sum.month.agg <- uza.weather[, .(min.temp.f = min(temp.f, na.rm = T),
                                   med.temp.f = median(temp.f, na.rm = T),
                                   max.temp.f = max(temp.f, na.rm = T)), by = .(hour,month,source)]
uza.weather.sum.month.agg <- uza.weather.sum.month.agg[month %in% c("Jun","Jul","Aug")]

sum.month.x.hour.p <- ggplot(uza.weather.sum.month.agg, aes(x = hour, y = med.temp.f , group = month)) +
  geom_ribbon(aes(ymin = min.temp.f, ymax = max.temp.f),  fill = "grey50", alpha = 0.5) +
  geom_line(size = 1.2) +
  geom_segment(aes(x = 0, y = 25, xend = 23, yend = 25)) +
  geom_segment(aes(x = 0, y = 25, xend = 0, yend = 135)) +
  facet_wrap(month ~ source, scales = "fixed") +
  labs(title = "Hourly Summer Temperatures Ranges for Phoenix UZA by Data Source, 2017", x = "Hour", y = "Temperature (deg F)") +
  scale_x_continuous(limits = c(0,23), breaks = c(0,6,12,18), expand = c(0,0)) +
  scale_y_continuous(limits = c(25,135), breaks = c(25,50,75,100,125), expand = c(0,0)) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", family = my.font, size = 12, colour = "black", hjust = 1),
        strip.text = element_text(family = my.font, size = 11, colour = "black"),
        axis.title = element_text(face = "bold", family = my.font, size = 12, colour = "black"),
        axis.text = element_text(family = my.font, size = 10, colour = "black"),
        axis.text.x = element_text(hjust = 0.25) # for some reason the text isn't defaulted to centered on the ticks
  
  )

sum.month.x.hour.p
#ggsave("hrly_tempF_by_sum_month_and_source_phx_UZA.png", sum.month.x.hour.p, device = "png", path = here("figures"), scale = 1, width = 6.5, height = 8, dpi = 300, units = "in")

# create summer data aggreate by station for GIF
uza.weather.summer.station.agg <- uza.weather[month %in% c("Jun","Jul","Aug"), .(min.temp.f = min(temp.f, na.rm = T),
                                   mean.temp.f = mean(temp.f, na.rm = T),
                                   max.temp.f = max(temp.f, na.rm = T)), by = .(hour,station.name)]

```

```{r analysis, include = FALSE}

# for each buffer distance around stations, calc the fraction of pavement and the total traffic

```

This document was compiled using R `r getRversion()`, R Markdown, and the following packages:  

```{r appendix, echo = FALSE, results = 'asis'}

# hide xtable global comment option
options(xtable.comment = FALSE)

# output table of packages used with packages listed in alphabetical order
print(xtable(package.info[order(package.info$Package),], row.names = NULL), include.rownames=FALSE)

```





