---
title: "Understanding the influence of transportation infrastrucure on Urban Heat in Phoenix, Arizona"
author: "C.G. Hoehne"
date: "October 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
# default to knitr option of code printing
knitr::opts_chunk$set(echo = TRUE)

# load fonts for use figures
extrafont::loadfonts(device = "win") 
##** CHANGE THIS TO LOAD ONLY THE FONT WE WILL USE
##** PREFERABLY INCLUDE THIS FONT IN THE REPO

# store master figure font as string
my.font <- "Century Gothic"

# list of all dependant packages
list.of.packages <- c("tidyverse",
                      "data.table", 
                      "lubridate",
                      "weathermetrics",
                      "rgdal",
                      "rgeos",
                      "raster",
                      "tmap",
                      "maptools",
                      "deldir",
                      "here")

# install missing packages
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

# load packages
lapply(list.of.packages, library, character.only = TRUE)

# store loaded packages and thier version in a data.frame
package.info <- data.frame("Package" = list.of.packages, "Version" = unlist(lapply(list.of.packages, function (x) as.character(packageVersion(x)))))

```

## Data Sources

### Weather Data
- National Centers for Environmental Information (NCEI): https://gis.ncdc.noaa.gov/maps/ncei/cdo/hourly
- The Arizona Meteorological Network (AZMET): https://cals.arizona.edu/azmet/az-data.htm
- Maricopa County Flood Control District (MCFCD): https://www.maricopa.gov/3769/Weather-Sensor-Data
- iButton data recorded in the Summer of 2017 by the Urban Water Innovation Network (UWIN): https://erams.com/UWIN/


### Parking Data
- Forthcoming publication by Hoehne et al, info here: http://transportationlca.org/phoenixparking/

### Road Network Data
- OpenStreetMap (© OpenStreetMap contributors)
- Arizona Department of Transportation (ADOT)

### Traffic Data
- Calibrated Maricopa Association of Governments (MAG) travel demand and parcel data via modifed version of SUMO (Simulation of Urban MObility) using OpenStreetMap street network data


```{r weather-data-prep, include = FALSE}

#############################################################################
## DATA IMPORT SCRIPT FOR 2017 PHOENIX, AZ WEATHER DATA - VARIOUS SOURCES ##
###########################################################################

#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#
# Import Arizona Meteorological Network data  #
#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#-#

encanto <- rbindlist(lapply(list.files(here(path = "data/AZMET/Encanto/raw"), pattern="txt$", full.names= T), read.csv, header = F))
desert.ridge <- rbindlist(lapply(list.files(here(path = "data/AZMET/Desert Ridge/raw"), pattern="txt$", full.names= T), read.csv, header = F))
mesa <- rbindlist(lapply(list.files(here(path = "data/AZMET/Mesa/raw"), pattern="txt$", full.names= T), read.csv, header = F))
greenway <- rbindlist(lapply(list.files(here(path = "data/AZMET/Greenway/raw"), pattern="txt$", full.names= T), read.csv, header = F))

# add station ID to each dataset
encanto$station <- "Encanto"
desert.ridge$station <- "Desert Ridge"
mesa$station <- "Mesa"
greenway$station <- "Greenway"

# bind all together, remove individual dfs
azmet.data <- rbind(encanto,desert.ridge,mesa,greenway)

# assign column names (see data dictionary)
colnames(azmet.data) <- c("year","day.year","hour","temp.c","rh","vap","sol","precip","soil.4.in","soil.20.in","winspd","wind.mag","windir","wind.dir.dev","max.wind.spd","ETo","act.vap","dewpt.c","station.name")

# fix temp = 999 to NA
azmet.data$temp.c <- ifelse(azmet.data$temp.c == 999, NA, azmet.data$temp.c)

# create fahrenheit temp/dew
azmet.data[, temp.f := ((temp.c * 1.8) + 32)]
azmet.data[, dewpt.f := ((dewpt.c * 1.8) + 32)]

# add date column
azmet.data$date <- as.Date(azmet.data$day.year-1, origin = paste0(azmet.data$year,"-01-01"))

# format date + time in new column as YYYY-MM-DD HH:MM:SS
azmet.data$date.time  <- ymd_hm(paste0(azmet.data$date," ",azmet.data$hour,":00"), tz = "US/Arizona")

# read AZMET station data
azmet.stations <- fread(here("data/AZMET/stations.csv"))

# add data source column and station.id column (NA, but for ordered cols later)
azmet.data$source <- "AZMET"
azmet.stations$source <- "AZMET"
azmet.data$station.id <- NA
azmet.stations$station.id <- NA

# keep relevant columns only
azmet.data <- azmet.data[,c("source","station.name","date.time","temp.f","dewpt.f","temp.c","dewpt.c","winspd","windir")]
azmet.stations <- azmet.stations[,c("source","station.name","lat","lon","elevation")]

# keep 2017 data only
azmet.data <- azmet.data[year(date.time) == 2017]


#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#
# Import National Centers for Environmental Information data  #
#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#~#

ncei.data <- read_fwf(here("data/NCEI/2016_data.txt"), skip = 1,
                       fwf_empty(here("data/NCEI/2016_data.txt"),
                                 col_names = c("station.id","WBAN","date","windir","winspd","GUS","CLG","SKC","L","M","H","VSB","MW1","MW2","MW3","MW4","AW1","AW2","AW3","AW4","W","temp.f","dewpt.f","SLP","ALT","STP","MAX","MIN","PCP01","PCP06","PCP24","PCPXX","SD")))

ncei.stations <- read_fwf(here(path = "data/NCEI/stations.txt"), skip = 2, fwf_widths(c(7,6,31,51,31,9,9,10), c("station.id","WBAN","station.name","COUNTRY","STATE","lat","lon","elevation")))

# set as data.table
setDT(ncei.data)
setDT(ncei.stations)

# convert variables to numeric (via characters because "NA" is "**")
ncei.data[, temp.f := as.numeric(as.character(temp.f))]
ncei.data[, dewpt.f := as.numeric(as.character(dewpt.f))]
ncei.data[, winspd := as.numeric(as.character(winspd))]
ncei.data[, windir := as.numeric(as.character(windir))]

# create celcius temp/dew
ncei.data[, temp.c := ((temp.f - 32) / 1.8)]
ncei.data[, dewpt.c := ((dewpt.f - 32) / 1.8)]

# format date + time in new column as YYYY-MM-DD HH:MM:SS
ncei.data$date.time  <- ymd_hm(ncei.data$date, tz = "UTC")

# convert to AZ timezone
ncei.data$date.time <- with_tz(ncei.data$date.time, tz = "US/Arizona") # change timezone to local (AZ)

# merge station.name to data
ncei.data <- merge(ncei.data, ncei.stations[,.(station.name,station.id)], by = "station.id")

# add data source column
ncei.data$source <- "NCEI"
ncei.stations$source <- "NCEI"

# keep relevant columns only
ncei.data <- ncei.data[,c("source","station.name","date.time","temp.f","dewpt.f","temp.c","dewpt.c","winspd","windir")]
ncei.stations <- ncei.stations[,c("source","station.name","lat","lon","elevation")]

# keep 2017 data only
ncei.data <- ncei.data[year(date.time) == 2017]


#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#
# Import Maricopa County Flood Control District data  #
#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#*#

# temp data
temp.files <- list.files(here(path = "data/MCFCD/2017/Temp"), pattern="txt$", full.names = T) # full file path names
temp.names <- gsub(".txt", "", list.files(here(path = "data/MCFCD/2017/Temp"), pattern = "txt$", full.names = F)) # names (stations) of files
mcfcd.temp.data <- rbindlist(lapply(temp.files, fread), idcol = "station.id") # load all station data 
mcfcd.temp.data[, station.id := factor(station.id, labels = basename(temp.names))] # add station names to column 'station'
mcfcd.temp.data[, station.id := as.character(station.id)] # convert to character for matching
colnames(mcfcd.temp.data) <- c("station.id","date","time","temp.f") # add rest of column names
mcfcd.temp.data$date.time  <- mdy_hms(paste0(mcfcd.temp.data$date," ",mcfcd.temp.data$time), tz = "US/Arizona") # MCFCD data is all local

# dewpt data
dewpt.files <- list.files(here(path = "data/MCFCD/2017/Dewpoint"), pattern="txt$", full.names = T) # full file path names
dewpt.names <- gsub(".txt", "", list.files(here(path = "data/MCFCD/2017/Dewpoint"), pattern = "txt$", full.names = F)) # names (stations) of files
mcfcd.dewpt.data <- rbindlist(lapply(dewpt.files, fread), idcol = "station.id") # load all station data 
mcfcd.dewpt.data[, station.id := factor(station.id, labels = basename(dewpt.names))] # add station names to column 'station'
mcfcd.dewpt.data[, station.id := as.character(station.id)] # convert to character for matching
colnames(mcfcd.dewpt.data) <- c("station.id","date","time","dewpt.f") # add rest of column names
mcfcd.dewpt.data$date.time  <- mdy_hms(paste0(mcfcd.dewpt.data$date," ",mcfcd.dewpt.data$time), tz = "US/Arizona") # MCFCD data is all local

# windir data
windir.files <- list.files(here(path = "data/MCFCD/2017/Wdir"), pattern="txt$", full.names = T) # full file path names
windir.names <- gsub(".txt", "", list.files(here(path = "data/MCFCD/2017/Wdir"), pattern = "txt$", full.names = F)) # names (stations) of files
mcfcd.windir.data <- rbindlist(lapply(windir.files, fread), idcol = "station.id") # load all station data 
mcfcd.windir.data[, station.id := factor(station.id, labels = basename(windir.names))] # add station names to column 'station'
mcfcd.windir.data[, station.id := as.character(station.id)] # convert to character for matching
colnames(mcfcd.windir.data) <- c("station.id","date","time","windir") # add rest of column names
mcfcd.windir.data$date.time  <- mdy_hms(paste0(mcfcd.windir.data$date," ",mcfcd.windir.data$time), tz = "US/Arizona") # MCFCD data is all local

# winspd data
winspd.files <- list.files(here(path = "data/MCFCD/2017/Wspd"), pattern="txt$", full.names = T) # full file path names
winspd.names <- gsub(".txt", "", list.files(here(path = "data/MCFCD/2017/Wspd"), pattern = "txt$", full.names = F)) # names (stations) of files
mcfcd.winspd.data <- rbindlist(lapply(winspd.files, fread), idcol = "station.id") # load all station data 
mcfcd.winspd.data[, station.id := factor(station.id, labels = basename(winspd.names))] # add station names to column 'station'
mcfcd.winspd.data[, station.id := as.character(station.id)] # convert to character for matching
colnames(mcfcd.winspd.data) <- c("station.id","date","time","winspd") # add rest of column names
mcfcd.winspd.data$date.time  <- mdy_hms(paste0(mcfcd.winspd.data$date," ",mcfcd.winspd.data$time), tz = "US/Arizona") # MCFCD data is all local

# solar data
solar.files <- list.files(here(path = "data/MCFCD/2017/Solar"), pattern="txt$", full.names = T) # full file path names
solar.names <- gsub(".txt", "", list.files(here(path = "data/MCFCD/2017/Solar"), pattern = "txt$", full.names = F)) # names (stations) of files
mcfcd.solar.data <- rbindlist(lapply(solar.files, fread), idcol = "station.id") # load all station data 
mcfcd.solar.data[, station.id := factor(station.id, labels = basename(solar.names))] # add station names to column 'station'
mcfcd.solar.data[, station.id := as.character(station.id)] # convert to character for matching
colnames(mcfcd.solar.data) <- c("station.id","date","time","solar") # add rest of column names
mcfcd.solar.data$date.time  <- mdy_hms(paste0(mcfcd.solar.data$date," ",mcfcd.solar.data$time), tz = "US/Arizona") # MCFCD data is all local

# station data
mcfcd.stations <- fread(here("data/MCFCD/stations.csv")) # load station data
mcfcd.stations <- mcfcd.stations[!is.na(station.id)]  # remove stations without an ID
mcfcd.stations[, station.id := as.character(station.id)] # convert station.id to character for matching
dms2dec <- function(x){ 
  z <- sapply(strsplit(x, " "), as.numeric)
  z[1, ] + z[2, ]/60 + z[3, ]/3600
}
mcfcd.stations$lat <- dms2dec(mcfcd.stations$lat.dms) # convert station deg-min-sec to decimal lat
mcfcd.stations$lon <- dms2dec(mcfcd.stations$lon.dms) # convert station deg-min-sec to decimal lon
mcfcd.stations$lon <- (-1) * mcfcd.stations$lon # arizona is west of prime meridian so longitude should be negative

# in temp data, stations 1001 and 28301 are missing (both outside urbanized area, so not essential to replace)
# in temp data, station 1002 is extra (removed, station is humidity data w/ 0 < x > 100)
# in dewpt data, stations 30816 and 34916 are extra (removed, no stations w/ those IDs in index, data is consistent with dewpt data tho)

# bind station names to each data type, then full join all data together by station.name and date.time 
#(note: solar rad data is only every 30 min instead of other four at 15 min)
mcfcd.temp.data <- merge(mcfcd.temp.data, mcfcd.stations[,.(station.name,station.id)], by = "station.id")
mcfcd.dewpt.data <- merge(mcfcd.dewpt.data, mcfcd.stations[,.(station.name,station.id)], by = "station.id")
mcfcd.windir.data <- merge(mcfcd.windir.data, mcfcd.stations[,.(station.name,station.id)], by = "station.id")
mcfcd.winspd.data <-merge(mcfcd.winspd.data, mcfcd.stations[,.(station.name,station.id)], by = "station.id")
mcfcd.solar.data <- merge(mcfcd.solar.data, mcfcd.stations[,.(station.name,station.id)], by = "station.id")

mcfcd.data <- merge(mcfcd.temp.data[,.(date.time,station.name,temp.f)], mcfcd.dewpt.data[,.(date.time,station.name,dewpt.f)], by = c("date.time","station.name"), all = T)
mcfcd.data <- merge(mcfcd.data, mcfcd.windir.data[,.(date.time,station.name,windir)], by = c("date.time","station.name"), all = T)
mcfcd.data <- merge(mcfcd.data, mcfcd.winspd.data[,.(date.time,station.name,winspd)], by = c("date.time","station.name"), all = T)
mcfcd.data <- merge(mcfcd.data, mcfcd.solar.data[,.(date.time,station.name,solar)], by = c("date.time","station.name"), all = T)

# create celcius temp/dew
mcfcd.data[, temp.c := ((temp.f - 32) / 1.8)]
mcfcd.data[, dewpt.c := ((dewpt.f - 32) / 1.8)]

# add data source column
mcfcd.data$source <- "MCFCD"
mcfcd.stations$source <- "MCFCD"

# filter out extraneous stations by sensor.type
mcfcd.stations <- mcfcd.stations[sensor.type %in% c("Temperature","Dewpoint","Wind Dir.","Peak Wind","Solar Rad.")]

# keep relevant columns only
mcfcd.stations <- mcfcd.stations[,c("source","station.name","lat","lon","elevation")]

# simplify station list
mcfcd.stations <- unique(mcfcd.stations)

# keep 2017 data only
mcfcd.data <- mcfcd.data[year(date.time) == 2017]


#+#+#+#+#+#+#+#+#+#+#+#+#+#+#+#+#+#+#
# Import UWIN Phoenix iButton data  #
#+#+#+#+#+#+#+#+#+#+#+#+#+#+#+#+#+#+#

# import temp and rel humidity data
ibut.temp <- fread(here("data/iButton/phxbuttons.csv"))
ibut.rhum <- fread(here("data/iButton/phxbuttonsrh.csv"))

# import station info and data
ibut.stations <- fread(here("data/iButton/phxdatsite_all.csv"))
ibut.stations.col.info <- fread(here("data/iButton/phxdatsiteall_col_description.csv"))

# rename some cols
setnames(ibut.temp, "Date - Time", "date.time") 
setnames(ibut.rhum, "Date - Time", "date.time")
setnames(ibut.stations, "ibuttonID", "station.name")
setnames(ibut.stations, "Lat", "lat")
setnames(ibut.stations, "Lon", "lon")

# keep only relevant columns
ibut.stations <- ibut.stations[, .(station.name,lat,lon)]

# remove hour column
ibut.temp[, hour := NULL] 
ibut.rhum[, hour := NULL] 

# melt and recast temp & hum data b/c in wide format
ibut.temp <- melt(ibut.temp, id = "date.time", variable.name = "station.name", value.name = "temp.c")
ibut.rhum <- melt(ibut.rhum, id = "date.time", variable.name = "station.name", value.name = "rhum.c")

# format date.time as YYYY-MM-DD HH:MM:SS
ibut.temp[, date.time := mdy_hm(date.time, tz = "US/Arizona")]
ibut.rhum[, date.time := mdy_hm(date.time, tz = "US/Arizona")]

# round times of temp cause they are 1 min shifted off
ibut.temp[, date.time := as.POSIXct(round.POSIXt(date.time, "hours"))]

# bind temp and hum toegether by station.id & date.time
ibut.data <- merge(ibut.temp, ibut.rhum, by = c("date.time","station.name"), all = T)

# convert rhum to dewpt
ibut.data[, dewpt.c := humidity.to.dewpoint(rh = rhum.c, t = temp.c, temperature.metric = "celsius")]
ibut.data[, rhum.c := NULL] 

# create fahrenheit temp/dew
ibut.data[, temp.f := ((temp.c * 1.8) + 32)]
ibut.data[, dewpt.f := ((dewpt.c * 1.8) + 32)]

# add data source column
ibut.data$source <- "iButton"
ibut.stations$source <- "iButton"


#^#^#^#^#^#^#^#^#^#^#^#^#^#
# Merge all data together #
#^#^#^#^#^#^#^#^#^#^#^#^#^#

# combine all data into one object, and all stations into one object
w.data <- rbindlist(list(azmet.data,ncei.data,mcfcd.data,ibut.data), use.names = T, fill = T)
w.stations <- rbindlist(list(azmet.stations,ncei.stations,mcfcd.stations,ibut.stations), use.names = T, fill = T)

# calculate heat index, (National Weather Surface improved estimate of Steadman eqn. (heat index calculator eqns)
w.data[, heat.f := heat.index(t = temp.f, dp = dewpt.f, temperature.metric = "fahrenheit", output.metric = "fahrenheit", round = 0)]
w.data[, heat.c := heat.index(t = temp.c, dp = dewpt.c, temperature.metric = "celsius", output.metric = "celsius", round = 1)]

# calculate observations in 2017 for each station by each data type
 for(station in w.stations$station.name){
   w.stations[station == station.name, n.temp := sum(!is.na(w.data$temp.f[station == w.data$station.name]))]
   w.stations[station == station.name, n.dewpt := sum(!is.na(w.data$dewpt.f[station == w.data$station.name]))]
   w.stations[station == station.name, n.windir := sum(!is.na(w.data$windir[station == w.data$station.name]))]
   w.stations[station == station.name, n.winspd := sum(!is.na(w.data$winspd[station == w.data$station.name]))]
   w.stations[station == station.name, n.solar := sum(!is.na(w.data$solar[station == w.data$station.name]))]
   w.stations[station == station.name, n.heat := sum(!is.na(w.data$heat.f[station == w.data$station.name]))]
 }

# remove stations with no temp obs
w.stations <- w.stations[n.temp != 0]

# remove everything but the formatted weather and station data
rm(list=setdiff(ls(), c("w.stations","w.data","package.info","my.font")))

```

## Plots

Here are the working plots:

```{r pressure, echo = FALSE}

# convert stations data.table w/ lat-lon to coordinates (SpatialPointDataFrame)
w.stations.spdf <- SpatialPointsDataFrame(coords = w.stations[, .(lon,lat)], data = w.stations,
                                          proj4string = CRS("+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"))

# various shapefiles for plotting
# load city labels shpfile
phx.labels <- shapefile(here("data/shapefiles/other/phx_metro_labels.shp"))

# load UZA & county border
uza.border <- shapefile(here("data/shapefiles/boundaries/maricopa_county_uza.shp"))
cnty.border <- shapefile(here("data/shapefiles/boundaries/maricopa_county.shp"))

# load 2017 travel ways (note these are cropped to extend just outside of the UZA)
hways <- shapefile(here("data/shapefiles/other/phx_metro_hways.shp"))

# transform stations crs to same as other shapefiles 
w.stations.spdf <- spTransform(w.stations.spdf, crs(uza.border))

# store palette for plotting stations
my.palette <- c("#0C120C","#640D14","#C20114","#6D7275")

# plot station points with UZA, highways, & city labels for context.
stations.plot <-   
  tm_shape(uza.border) + tm_borders(lwd = 0.8, lty = "solid", col = "grey40", alpha = 0.7) + tm_fill(col = "grey90") + # urbanized area (UZA) border & fill
  tm_shape(w.stations.spdf, bbox = w.stations.spdf) + tm_dots(col = "source", border.col = NULL, palette = my.palette, size = 0.2, alpha = 0.8, title = "Data Source") + # station points
  tm_shape(cnty.border) + tm_borders(lwd = 0.8, lty = "solid", col = "grey20", alpha = 0.7) + # county border
  tm_scale_bar(position = c(0.4,0.0), breaks = c(0,5,10,15,20), size = 0.90, color.light = "grey85") + # scalebar
  tm_compass(north = 0, type = "4star", size = 2, show.labels = 1, position = c(0.9,0.85)) + # compass
  tm_shape(hways) + tm_lines(lwd = 1.1, col = "grey40") + # highways
  tm_shape(phx.labels) + tm_text("name", size = .5, fontfamily = my.font,  fontface = "bold.italic", just = "center") + # city labels
  tm_layout(fontfamily = my.font, fontface = "italic", bg.color = "grey95", title.size = 1.5, legend.title.size = 1.10, # theme & formatting
            legend.text.size = 0.95, title = "", title.position = c(0.12,0.94),
            legend.position = c(0.02,0.45), outer.margins = c(0,0,0,0), asp = 0)

#tmap_save(stations.plot, filename = here("figures/stations_w_ibut.png"))
stations.plot


# calculate range of temps and mean temp for urbanized stations

# first clip stations points to ones w/in uza 
uza.stations <- raster::intersect(w.stations.spdf, uza.border)

# filter station list to uza stations
uza.stations.list <- w.stations[station.name %in% uza.stations$station.name]

# filter data to uza.station.list
uza.data <- w.data[station.name %in% uza.stations.list$station.name]

# if time is within 15 min of hour, round to hour, otherwise drop, then
# aggregate to the min, mean, max temp at each unique timestep for all stations
uza.data[, date.time.round := as.POSIXct(round.POSIXt(date.time, "hours"))]
uza.data <- uza.data[abs(difftime(date.time, date.time.round, units = "min")) < 15]
uza.data.agg <- uza.data[, .(min.temp.f = min(temp.f, na.rm = T),
                             med.temp.f = median(temp.f, na.rm = T),
                             max.temp.f = max(temp.f, na.rm = T)), by = .(date.time.round)]

# hourly temp min/mean/max by 2017 month aggregate across all stations
uza.data[, month := as.factor(month(date.time.round, label = T))]
uza.data[, hour := hour(date.time.round)]
uza.data.month.agg <- uza.data[, .(min.temp.f = min(temp.f, na.rm = T),
                                   med.temp.f = median(temp.f, na.rm = T),
                                   max.temp.f = max(temp.f, na.rm = T)), by = .(hour,month)]

# plot hourly temp.f range by month for 2017 for all stations
# Plot historical tempeature plots of metro regions
month.x.hour.p <- ggplot(uza.data.month.agg, aes(x = hour, y = med.temp.f , group = month)) +
  geom_ribbon(aes(ymin = min.temp.f, ymax = max.temp.f),  fill = "grey50", alpha = 0.5) +
  geom_line(size = 1.2) +
  geom_segment(aes(x = 0, y = 25, xend = 23, yend = 25)) +
  geom_segment(aes(x = 0, y = 25, xend = 0, yend = 135)) +
  facet_wrap(~ month , scales = "fixed") +
  labs(title = "Hourly Temperature Ranges by month for Phoenix UZA, 2017", x = "Hour", y = "Temperature (deg F)") +
  scale_x_continuous(limits = c(0,23), breaks = c(0,6,12,18), expand = c(0,0)) +
  scale_y_continuous(limits = c(25,135), breaks = c(25,50,75,100,125), expand = c(0,0)) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", family = my.font, size = 12, colour = "black", hjust = 1),
        strip.text = element_text(family = my.font, size = 11, colour = "black"),
        axis.title = element_text(face = "bold", family = my.font, size = 12, colour = "black"),
        axis.text = element_text(family = my.font, size = 10, colour = "black"),
        axis.text.x = element_text(hjust = 0.25) # for some reason the text isn't defaulted to centered on the ticks
  )

month.x.hour.p
#ggsave("hrly_tempF_by_month_all_phx_UZA_stations.png", month.x.hour.p, device = "png", path = here("figures"), scale = 1, width = 6.5, height = 8, dpi = 300, units = "in")

# plot hourly temp.f range for summer months for 2017 for all stations
uza.data.sum.month.agg <- uza.data[, .(min.temp.f = min(temp.f, na.rm = T),
                                   med.temp.f = median(temp.f, na.rm = T),
                                   max.temp.f = max(temp.f, na.rm = T)), by = .(hour,month,source)]
uza.data.sum.month.agg <- uza.data.sum.month.agg[month %in% c("Jun","Jul","Aug")]

sum.month.x.hour.p <- ggplot(uza.data.sum.month.agg, aes(x = hour, y = med.temp.f , group = month)) +
  geom_ribbon(aes(ymin = min.temp.f, ymax = max.temp.f),  fill = "grey50", alpha = 0.5) +
  geom_line(size = 1.2) +
  geom_segment(aes(x = 0, y = 25, xend = 23, yend = 25)) +
  geom_segment(aes(x = 0, y = 25, xend = 0, yend = 135)) +
  facet_wrap(month ~ source, scales = "fixed") +
  labs(title = "Hourly Summer Temperatures Ranges for Phoenix UZA by Data Source, 2017", x = "Hour", y = "Temperature (deg F)") +
  scale_x_continuous(limits = c(0,23), breaks = c(0,6,12,18), expand = c(0,0)) +
  scale_y_continuous(limits = c(25,135), breaks = c(25,50,75,100,125), expand = c(0,0)) +
  theme_minimal() +
  theme(plot.title = element_text(face = "bold", family = my.font, size = 12, colour = "black", hjust = 1),
        strip.text = element_text(family = my.font, size = 11, colour = "black"),
        axis.title = element_text(face = "bold", family = my.font, size = 12, colour = "black"),
        axis.text = element_text(family = my.font, size = 10, colour = "black"),
        axis.text.x = element_text(hjust = 0.25) # for some reason the text isn't defaulted to centered on the ticks
  
  )

sum.month.x.hour.p
#ggsave("hrly_tempF_by_sum_month_and_source_phx_UZA.png", sum.month.x.hour.p, device = "png", path = here("figures"), scale = 1, width = 6.5, height = 8, dpi = 300, units = "in")


# create summer data aggreate by station for GIF
uza.data.summer.station.agg <- uza.data[month %in% c("Jun","Jul","Aug"), .(min.temp.f = min(temp.f, na.rm = T),
                                   mean.temp.f = mean(temp.f, na.rm = T),
                                   max.temp.f = max(temp.f, na.rm = T)), by = .(hour,station.name)]
# save data for GIF



```

This document was compiled using R `r getRversion()`, R Markdown, and the following packages:
```{r package-table, echo = FALSE}
knitr::kable(package.info)
```





